---
title: "Bandit study group"
author: "Eric B. Laber"
date: "May 22, 2016"
output: html_document
---

###Basic bandit simulations
We consider a bandit algorithm with $K$ arms labeled $1,\ldots, K$.  First, 
lets compute the pseudo-regret for uniform guessing, i.e., $\left\lbrace I_t 
\right\rbrace_{t\ge 1}$ are i.i.d. with $I_t \sim
\mathrm{Uniform}\left\lbrace 1,\ldots, K\right\rbrace$. 

```{r}
library (ggplot2);
### We'll assume K=5 arms each with a beta-distn 
K = 5;
alphaSeq = c(5, 4.85, 4.5, 4.5, 1);
betaSeq = c(1, 1.2, 1.5, 3, 1);
muStar= 5/6;  # mean is alpha/(alpha + beta)
nSteps = 100;
regret = numeric (nSteps);
arms = numeric (nSteps);
for (i in 1:nSteps){
  arm = sample (K, size=1);
  arms[i] = arm;
  regret[i] = muStar - rbeta (1, shape1=alphaSeq[arm], shape2=betaSeq[arm]);
}

group = rep(1:K, nSteps);
val = numeric(K*nSteps);
tm = numeric (K*nSteps);
for (j in 1:K){
  armTotal = cumsum(arms==j);
  for (i in 1:nSteps){
    ind = (i-1)*K + j;
    val[ind] = armTotal[i];
    tm[ind] = i;
  }
}
plotDF = data.frame(time=tm, val=val, group=group);

plot(cumsum(regret), type='l', xlab="steps", ylab="regret");
ggplot (plotDF, aes(x=time, y=val, group=group, fill=group)) + geom_area(position="fill");


```



For our second bandit example, we'll sample each arm $m$ times and then choose
the argmax thereafter.  


```{r}
K = 5;
alphaSeq = c(5, 4.85, 4.5, 4.5, 1);
betaSeq = c(1, 1.2, 1.5, 3, 1);
muStar= 5/6;  # mean is alpha/(alpha + beta)
m = 15;
nSteps = 250;
regret = numeric (nSteps);
arms = numeric (nSteps);
muHat = numeric (K);
for (i in 1:(m*K)){
  arm = (i %% K) + 1;
  arms[i] = arm;
  xt  = rbeta (1, shape1=alphaSeq[arm], shape2=betaSeq[arm]);
  muHat[arm] = muHat[arm] + xt;
  regret[i] = muStar - xt;
}

armHat = which.max(muHat);
for (i in (m*K+1):nSteps){
  arm = armHat;  # constant 
  arms[i] = arm;
  regret[i] = muStar - rbeta (1, shape1=alphaSeq[arm], shape2=betaSeq[arm]);
}


group = rep(1:K, nSteps);
val = numeric(K*nSteps);
tm = numeric (K*nSteps);
for (j in 1:K){
  armTotal = cumsum(arms==j);
  for (i in 1:nSteps){
    ind = (i-1)*K + j;
    val[ind] = armTotal[i];
    tm[ind] = i;
  }
}
plotDF = data.frame(time=tm, val=val, group=group);

plot(cumsum(regret), type='l', xlab="steps", ylab="regeret");
ggplot (plotDF, aes(x=time, y=val, group=group, fill=group)) + geom_area(position="fill");

```


Now let's apply UCB with $\alpha=2.5$ and $\psi(\lambda) = \lambda^2/8$.  


```{r}
K = 5;
alphaSeq = c(5, 4.85, 4.5, 4.5, 1);
betaSeq = c(1, 1.2, 1.5, 3, 1);
muStar= 5/6;  # mean is alpha/(alpha + beta)
nSteps = 5000;
regret = numeric (nSteps);
arms = numeric (nSteps);
muSum = numeric (K);
nt = numeric (K);
for (i in 1:nSteps){
  mt = which(nt == 0);
  if (length(mt) > 0){
    arm = sample (mt, size=1);
  } else{
    aMax = which.max (muSum/nt + sqrt(20*log(i)/nt));
    arm = sample (aMax, size=1);
  }
  
  nt[arm] = nt[arm] + 1;
  arms[i] = arm;
  xt = rbeta (1, shape1=alphaSeq[arm], shape2=betaSeq[arm]);
  muSum[arm] = muSum[arm] + xt; 
  regret[i] = muStar - xt;
  }

group = rep(1:K, nSteps);
val = numeric(K*nSteps);
tm = numeric (K*nSteps);
for (j in 1:K){
  armTotal = cumsum(arms==j);
  for (i in 1:nSteps){
    ind = (i-1)*K + j;
    val[ind] = armTotal[i];
    tm[ind] = i;
  }
}
plotDF = data.frame(time=tm, val=val, group=group);

plot(cumsum(regret), type='l', xlab="steps", ylab="regeret");
ggplot (plotDF, aes(x=time, y=val, group=group, fill=group)) + geom_area(position="fill");


```



